# Connecting to MongoDB

## Mongo URI Format

The `mongo()` function initiates a connection object to a MongoDB server. For example:

```{r}
library(mongolite)
m <- mongo("mtcars", url = "mongodb://readwrite:test@ds043942.mongolab.com:43942/jeroen_test")
```

To get an overview of available methods, simply print the object to the terminal. 

```{r}
print(m)
```

The R manual page for the `mongo()` function gives some brief descriptions as well.

```r
?mongo
```

The manual page tells us that `mongo()` supports the following arguments:

 - `collection`: name of the collection to connect to. Defaults to `"test"`.
 - `db`: name of the database to connect to. Defaults to `"test"`.
 - `url`: address of the MongoDB server in standard [URI Format](https://docs.mongodb.com/manual/reference/connection-string/).
 - `verbose`: if `TRUE`, emits some extra output
 - `options`: additinoal connection options such as SSL keys/certs.
 
The `url` parameter contains a special URI format which defines the server address
and additional connection options. 

```
mongodb://[username:password@]host1[:port1][,host2[:port2],...[/[database][?options]]
```

The [Mongo Connection String Manual](https://docs.mongodb.com/manual/reference/connection-string/)
gives an overview of the connection string syntax and options. 
Below the most important options for using mongolite.

### DNS Seedlist Connection Format

New in `mongolite 1.3` is support for [seedlist](https://docs.mongodb.com/manual/reference/connection-string/#dns-seedlist-connection-format) URLs with the `mongodb+srv://` prefix. This indicates that before connecting, the client should lookup the actual host addresses and parameters from the DNS [SRV](https://en.wikipedia.org/wiki/SRV_record) or [TXT](https://en.wikipedia.org/wiki/TXT_record) record.

```{r}
con <- mongo("mtcars", url = "mongodb+srv://readwrite:test@cluster0-84vdt.mongodb.net/test?authSource=admin")
con$insert(mtcars)
con$drop()
```

The DNS seedlist allows for using a short and fixed URL for clusters consisting of multiple or dynamic servers and parameters.

## Authentication

MongoDB supports several authentication modes.

### LDAP

```{r, eval = FALSE}
USER = "drivers-team"
PASS = "mongor0x$xgen"
HOST = "ldaptest.10gen.cc"

# Using plain-text
URI = sprintf("mongodb://%s:%s@%s/ldap?authMechanism=PLAIN", USER, PASS, HOST)
m <- mongo(url = URI)
m$find()
```

However it is recommended to use SSL instead of plain text when authenticating with a username/password: 

```{r, eval = FALSE}
# Or over SSL
m <- mongo(url = paste0(URI, "&ssl=true"), options = 
             ssl_options(ca = "auth/ca.crt", allow_invalid_hostname = TRUE))
m$find()
```

### X509

Let's check if our server supports SSL:

```{r, eval = FALSE}
certs <- openssl::download_ssl_cert('ldaptest.10gen.cc', 27017)
print(certs)
str(as.list(certs[[1]]))
```

To use X509 authentication the Mongo URI needs `ssl=true&authMechanism=MONGODB-X509`:

```{r, eval = FALSE}
# Using X509 SSL auth
HOST <- "ldaptest.10gen.cc"
USER <- "CN=client,OU=kerneluser,O=10Gen,L=New York City,ST=New York,C=US"
URI <- sprintf("mongodb://%s@%s/x509?ssl=true&authMechanism=MONGODB-X509", USER, HOST)
OPTS <- ssl_options(cert = "auth/client.pem", key = "auth/key.pem", ca = "auth/ca.crt", allow_invalid_hostname = TRUE)
m <- mongo(url = URI, options = OPTS)
m$find()
```

### Kerberos

*__Note__: Windows uses `SSPI` for Kerberos authentication. This section does not apply*.  

Kerberos authentication on Linux requires installation of a Kerberos client. 
On OS-X Kerberos is already installed by default. 
On Ubuntu/Debian we need:

```
sudo apt-get install krb5-user libsasl2-modules-gssapi-mit
```

Next, create or edit `/etc/krbs5.conf` and add our server under `[realms]` for example:

```
[realms]
  LDAPTEST.10GEN.CC = {
    kdc = ldaptest.10gen.cc
    admin_server = ldaptest.10gen.cc
  }
```

In a terminal run the following (only have to do this once)

```
kinit drivers@LDAPTEST.10GEN.CC
klist
```

We should now be able to connect in R:

```{r, eval = FALSE}
HOST <- "ldaptest.10gen.cc"
USER <- "drivers%40LDAPTEST.10GEN.CC"
URI <- sprintf("mongodb://%s@%s/kerberos?authMechanism=GSSAPI", USER, HOST)
m <- mongo(url = URI)
m$find()
```


## SSH Tunnel

To connect to MongoDB via an SSH tunnel, you need to setup the tunnel separately with an SSH client. For example the mongolite manual contains this example:

```r
con <- mongo("mtcars", url = "mongodb://readwrite:test@ds043942.mongolab.com:43942/jeroen_test")
```

Assume we want to tunnel through `dev.opencpu.org` which runs an SSH server on the standard port 22 with username `jeroen`. To initiate a tunnel from `localhost:9999` to `ds043942.mongolab.com:43942` via the ssh server `dev.opencpu.org:22`, open a terminal and run:

```
ssh -L 9999:ds043942.mongolab.com:43942 jeroen@dev.opencpu.org -vN -p22
```

Some relevant `ssh` flags:

 - `-v` (optional) show verbose status output
 - `-f` run the tunnel server in the background. Use `pkill ssh` to kill.
 - `-p22` connect to ssh server on port 22 (default)
 - `-i/some/path/id_rsa` authenticate with ssh using a private key

Check `man ssh` for more ssh options It is also possible to run this command directly from R:

```{r}
system("ssh -L 9999:ds043942.mongolab.com:43942 jeroen@dev.opencpu.org -fN -p22")
```

Once tunnel has been established, we can connect to our our ssh client which will tunnel traffic to our MongoDB server. In our example we run the ssh client on our localhost port 9999:

```{r}
con <- mongo("mtcars", url = "mongodb://readwrite:test@localhost:9999/jeroen_test")
con$insert(mtcars)
```

```{r echo = FALSE}
con$drop()
system("pkill ssh")
```

If you want to setup a tunnel client on Windows and you do not have the `ssh` program, you can an SSH client like putty to setup the tunnel. See [this example](http://realprogrammers.com/how_to/set_up_an_ssh_tunnel_with_putty.html).


## SSL options

For security reasons, SSL options can not be configured in the URI but have to be set 
manually via the `options` parameter. The `ssl_options` function shows the default values:

```{r}
ssl_options()
```

You can use this function to specify connection SSL options:

```r
m <- mongo(url = "mongodb://localhost?ssl=true", 
  options = ssl_options(cert = "~/client.crt", key = "~/id_rsa.pem"))
```

The MongoDB [SSL client manual](https://docs.mongodb.com/manual/tutorial/configure-ssl-clients/)
has more detailed descriptions on the various options.

## Replica Options

The URI accepts a few special keys when connecting to a replicaset. The
[connection-string manual](https://docs.mongodb.com/manual/reference/connection-string/)
is the canonical source for all parameters. Most users should stick with the
defaults here, only specify these if you know what you are doing.


### Read Preference

The **Read Preference** paremeter specifies if the client should connect to the 
primary node (default) or a secondary node in the replica set. 

```r
m <- mongo(url = "mongodb://host.example.com/?readPreference=secondary")
```

### Write Concern

The **Write Concern** parameter is used to specify the level of acknowledgement that 
the write operation has propagated to a number of server nodes. The url string parameter
is the letter `w`.

```r
m <- mongo(url = "mongodb://host.example.com/?w=2")
```

Note that specifying this parameter to 2 on a server that is not a replicaset will result in an error when trying to write:

```{r, error=TRUE}
m <- mongo(url = "mongodb://localhost/?w=2")
m$insert('{"foo" : "bar"}')
```

### Read Concern

Finally, **Read Concern** allows clients to choose a level of isolation for their reads from replica sets. The default value `local` returns the instance’s most recent data, but provides no guarantee that the data has been written to a majority of the replica set members (i.e. may be rolled back).

On the other hand, if we specify `majority` the server will only return data that has been propagated to the majority of nodes.

```{r, error=TRUE}
m <- mongo(url = "mongodb://localhost/?readConcernLevel=majority")
m$insert('{"foo": 123}')
m$insert('{"foo": 456}')
```

In the case of our local single-node server this is never the case. Therefore we see that the server does not return any data that meets the majority level.

```{r, error=TRUE}
m$count()
m$find()
```

The data is definitely there though, it just doesn't meet the `majority` criterium. If we create a new connection with level `local` we do get to see our data:

```{r}
m2 <- mongo(url = "mongodb://localhost/?readConcernLevel=local")
m2$count()
m2$find()
```

## Global options

Finally the `mongo_options` method allows for setting global client options that span across connections. Currently two options are supported:

 - `log_level` set the mongo [log level](http://mongoc.org/libmongoc/current/logging.html), e.g. for printing debugging information.
 - `bigint_as_char` set to `TRUE` to parse int64 numbers as strings rather than doubles (R does not support large integers natively)
 
The default values are:

```{r}
mongo_options()
```
 
See the manual page for `?mongo_options` for more details.
