# Connecting to MongoDB

## Mongo URI Format

The `mongo()` function initiates a connection object to a MongoDB server. For example:

```{r}
library(mongolite)
m <- mongo("mtcars", url = "mongodb://readwrite:test@ds043942.mongolab.com:43942/jeroen_test")
```

To get an overview of available methods, simply print the object to the terminal. 

```{r}
print(m)
```

The R manual page for the `mongo()` function gives some brief descriptions as well.

```r
?mongo
```

The manual page tells us that `mongo()` supports the following arguments:

 - `collection`: name of the collection to connect to. Defaults to `"test"`.
 - `db`: name of the database to connect to. Defaults to `"test"`.
 - `url`: address of the MongoDB server in standard [URI Format](https://docs.mongodb.com/manual/reference/connection-string/).
 - `verbose`: if `TRUE`, emits some extra output
 - `options`: additinoal connection options such as SSL keys/certs.
 
The `url` parameter contains a special URI format which defines the server address
and additional connection options. 

```
mongodb://[username:password@]host1[:port1][,host2[:port2],...[/[database][?options]]
```

The [Mongo Connection String Manual](https://docs.mongodb.com/manual/reference/connection-string/)
gives an overview of the connection string syntax and options. 
Below the most important options for using mongolite.



## Authentication

MongoDB supports several authentication modes.

### LDAP

```{r, eval = FALSE}
USER = "drivers-team"
PASS = "mongor0x$xgen"
HOST = "ldaptest.10gen.cc"

# Using plain-text
URI = sprintf("mongodb://%s:%s@%s/ldap?authMechanism=PLAIN", USER, PASS, HOST)
m <- mongo(url = URI)
m$find()
```

However it is recommended to use SSL instead of plain text when authenticating with a username/password: 

```{r, eval = FALSE}
# Or over SSL
m <- mongo(url = paste0(URI, "&ssl=true"), options = 
             ssl_options(ca = "auth/ca.crt", allow_invalid_hostname = TRUE))
m$find()
```

### X509

Let's check if our server supports SSL:

```{r, eval = FALSE}
certs <- openssl::download_ssl_cert('ldaptest.10gen.cc', 27017)
print(certs)
str(as.list(certs[[1]]))
```

To use X509 authentication the Mongo URI needs `ssl=true&authMechanism=MONGODB-X509`:

```{r, eval = FALSE}
# Using X509 SSL auth
HOST <- "ldaptest.10gen.cc"
USER <- "CN=client,OU=kerneluser,O=10Gen,L=New York City,ST=New York,C=US"
URI <- sprintf("mongodb://%s@%s/x509?ssl=true&authMechanism=MONGODB-X509", USER, HOST)
OPTS <- ssl_options(cert = "auth/client.pem", key = "auth/key.pem", ca = "auth/ca.crt", allow_invalid_hostname = TRUE)
m <- mongo(url = URI, options = OPTS)
m$find()
```

### Kerberos

*__Note__: Windows uses `SSPI` for Kerberos authentication. This section does not apply*.  

Kerberos authentication on Linux requires installation of a Kerberos client. 
On OS-X Kerberos is already installed by default. 
On Ubuntu/Debian we need:

```
sudo apt-get install krb5-user libsasl2-modules-gssapi-mit
```

Next, create or edit `/etc/krbs5.conf` and add our server under `[realms]` for example:

```
[realms]
  LDAPTEST.10GEN.CC = {
    kdc = ldaptest.10gen.cc
    admin_server = ldaptest.10gen.cc
  }
```

In a terminal run the following (only have to do this once)

```
kinit drivers@LDAPTEST.10GEN.CC
klist
```

We should now be able to connect in R:

```{r, eval = FALSE}
HOST <- "ldaptest.10gen.cc"
USER <- "drivers%40LDAPTEST.10GEN.CC"
URI <- sprintf("mongodb://%s@%s/kerberos?authMechanism=GSSAPI", USER, HOST)
m <- mongo(url = URI)
m$find()
```


## SSL options

For security reasons, SSL options can not be configured in the URI but have to be set 
manually via the `options` parameter. The `ssl_options` function shows the default values:

```{r}
ssl_options()
```

You can use this function to specify connection SSL options:

```r
m <- mongo(url = "mongodb://localhost?ssl=true", 
  options = ssl_options(cert = "~/client.crt", key = "~/id_rsa.pem"))
```

The MongoDB [SSL client manual](https://docs.mongodb.com/manual/tutorial/configure-ssl-clients/)
has more detailed descriptions on the various options.

## Replica Options

The URI accepts a few special keys when connecting to a replicaset. The
[connection-string manual](https://docs.mongodb.com/manual/reference/connection-string/)
is the canonical source for all parameters. Most users should stick with the
defaults here, only specify these if you know what you are doing.


### Read Preference

The **Read Preference** paremeter specifies if the client should connect to the 
primary node (default) or a secondary node in the replica set. 

```r
m <- mongo(url = "mongodb://host.example.com/?readPreference=secondary")
```

### Write Concern

The **Write Concern** parameter is used to specify the level of acknowledgement that 
the write operation has propagated to a number of server nodes. The url string parameter
is the letter `w`.

```r
m <- mongo(url = "mongodb://host.example.com/?w=2")
```

Note that specifying this parameter to 2 on a server that is not a replicaset will result in an error when trying to write:

```{r, error=TRUE}
m <- mongo(url = "mongodb://localhost/?w=2")
m$insert('{"foo" : "bar"}')
```

### Read Concern

Finally, **Read Concern** allows clients to choose a level of isolation for their reads from replica sets. The default value `local` returns the instance’s most recent data, but provides no guarantee that the data has been written to a majority of the replica set members (i.e. may be rolled back).

On the other hand, if we specify `majority` the server will only return data that has been propagated to the majority of nodes.

```{r, error=TRUE}
m <- mongo(url = "mongodb://localhost/?readConcernLevel=majority")
m$insert('{"foo": 123}')
m$insert('{"foo": 456}')
```

In the case of our local single-node server this is never the case. Therefore we see that the server does not return any data that meets the majority level.

```{r, error=TRUE}
m$count()
m$find()
```

The data is definitely there though, it just doesn't meet the `majority` criterium. If we create a new connection with level `local` we do get to see our data:

```{r}
m2 <- mongo(url = "mongodb://localhost/?readConcernLevel=local")
m2$count()
m2$find()
```

## Global options

Finally the `mongo_options` method allows for setting global client options that span across connections. Currently two options are supported:

 - `log_level` set the mongo [log level](http://mongoc.org/libmongoc/current/logging.html), e.g. for printing debugging information.
 - `bigint_as_char` set to `TRUE` to parse int64 numbers as strings rather than doubles (R does not support large integers natively)
 
The default values are:

```{r}
mongo_options()
```
 
See the manual page for `?mongo_options` for more details.
